Stack Safety is a Security Property

### 2021-04-11 tex/main.tex
Tweak to lazy description
Redid Assumptions/Threat model section

To illustrate the applicability of these properties, we perform randomized testing to
validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces
them. Randomized testing is a relatively low-cost way to validate a potential enforcement
mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
covers potentially erroneous programs, we test intentionally buggy policies alongside
the real one. If the testing can differentiate correct implementations from buggy ones,
it is an argument in favor of their correctness.
%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
%  significant contribution, and partly to give the reader some context for
%  why this is part of our story.  E.g., ``One nice thing we can do with our
%  spec is to use it to evaluate the claims of existing or future proposals
%  for enforcement mechanisms.  One relatively inexpensive way to carry out
%  such an evaluation is to see whether property-based testing against this
%  spec can distinguish correct implementations of the mechanism from buggy
%  ones...''}
Of course, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
Clearing micro-policy does not enforce stepwise integrity.
Unfortunately the existing Lazy Tagging and Clearing
micro-policy does not quite enforce observable confidentiality either. %, as we show in
%testing~\apt{really??},
but formalizing the ideal property allows us to see how it falls short.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
observational integrity, at some performance cost.\apt{but we're not testing that, right?}
\rb{Not sure about the property part right now, the machine side of this is
  pretty much done}
\sna{It's aspirational so far, but there's no fundamental reason we can't test it. We may
  need a caveat about how it's harder to test?}

@@

We begin with a description of our threat model (\cref{sec:threat}),
then show how our properties apply to examples of stack unsafety 
(\cref{sec:running-example}). Then, after some technical preliminaries
(\cref{sec:prelim}), we formalize the properties (\cref{sec:eager}, \cref{sec:lazy})
and describe testing (\cref{sec:testing}).

\section{Threat Model and Assumptions}
\label{sec:threat}

@@

Intuititively, stack safety means that a caller's stack frame is protected
from a confused or malicious callee's
attempts to read or write its private data or hijack its control flow. And in fact the caller,
while not malicious, may itself be confused or erroneous. Our attacker
model reflects this intention\bcp{I don't understand how it reflects this
  intuition about the caller---indeed, I don't understand the point about
  the caller.} by assuming that both the attacker and the defender
may run arbitrary machine code. Hardware and timing attacks are out of scope.

@@

In this very strong attacker model, there is little obvious distinction between
the attacker and the defender. They are separated only by the control transfer
inherent in their calls and returns. Our stack-safety concepts therefore depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code (the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes).
%
We assume we are given a machine language program annotated with the locations of
instructions representing calls, returns, and function entry points. Although this
information would typically come from a compiler together with the machine code,
we do not assume anything about their provenance. This means that attackers (or
foolish defenders) are under no obligation to have reasonable annotations.

@@

Since we rely so heavily on code annotations, we must ask: what if the annotations
are wrong? First, suppose a call or return occurs that is not annotated.
In this case, the answer is simple: as far as the properties are concerned, there has
been no call or return, only an instruction sequence resembling such an operation, and
therefore no associated change in permissions. This is particularly interesting in the
case of a false return, where control may appear to have returned to the caller, but
the enforcement mechanism must still protect the caller's memory.

@@

\paragraph*{System Assumptions}

Our properties are defined in terms of an abstract machine model that makes very few
assumptions about the machine that instantiates it.  We assume a register
machine with a program counter \(\PCname\) and stack pointer \(\rsp\); other
details of the machine's ISA are unimportant. We assume that we know ahead of time
the layout of memory, including code (with annotations) and which addresses are within
the bounds of the stack.
%
We further parameterize over the structure of the stack, particularly the means by
which a caller identifies the data that should be protected during a call. We term
this the {\em sealing convention} and model it as a relation between machine states
and addresses. Then from any call, those the sealing convention relates to the
state will be protected until the return.

\paragraph*{Policies}

@@

\paragraph*{Micro-Policies}
%
Micro-Policies are a flexible tag-based, hardware-accelerated
reference monitoring mechanism; they been applied to stack safety
enforcement by \citet{DBLP:conf/sp/RoesslerD18}.
Here the policy state consists of metadata tags (e.g., identifying stack frames)
attached to each value in memory or registers, and the policy step
function checks that each machine operation obeys a set of rules on tags
(e.g., that the current $\PCname$ tag matches the tag on the stack
location being accessed), and halts the machine if not.  Distinguishing
ordinary machine state and policy state is very natural in this setting.
%% Here the machine state is extended with tags paired with all components of the
%% system, and a monitor checks for each operation that the tags on the operands
%% obey a set of rules. One could treat tags directly as part of the machine
%% state, and build the rules into the step function. But local state
%% encapsulation properties as we frame them are not intuitive in their
%% interaction with tags or other additional enforcement state, so it is helpful
%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
%% a software fail state.
%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
%%   policy? }

\ifaftersubmission\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
%These approaches are represented as policies that
%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
%being produced by the compiler.) The set of policy states is again the unit type
%and the step function constant. Property enforcement comes from the fact that
%only ill-formed states can induce a trace that violates the property.

>>>>

To illustrate the applicability of these properties, we perform randomized testing to
validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces
them. Randomized testing is a relatively low-cost way to validate a potential enforcement
mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
covers potentially erroneous programs, we test intentionally buggy policies alongside
the real one. If the testing can differentiate correct implementations from buggy ones,
it is an argument in favor of their correctness.
%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
%  significant contribution, and partly to give the reader some context for
%  why this is part of our story.  E.g., ``One nice thing we can do with our
%  spec is to use it to evaluate the claims of existing or future proposals
%  for enforcement mechanisms.  One relatively inexpensive way to carry out
%  such an evaluation is to see whether property-based testing against this
%  spec can distinguish correct implementations of the mechanism from buggy
%  ones...''}
Of course, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
Clearing micro-policy does not enforce stepwise integrity.
Unfortunately the existing Lazy Tagging and Clearing
micro-policy does not quite enforce observable confidentiality either. %, as we show in
%testing~\apt{really??},
But formalizing the ideal property allows us to see how it falls short -- a specific
violation of the temporal aspect of confidentiality.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
observational integrity, at some performance cost.\apt{but we're not testing that, right?}
\rb{Not sure about the property part right now, the machine side of this is
  pretty much done}
\sna{It's aspirational so far, but there's no fundamental reason we can't test it. We may
  need a caveat about how it's harder to test?}

@@

We begin with a description of our assumptions and threat model (\cref{sec:threat}),
then show how our properties apply to examples of stack unsafety 
(\cref{sec:running-example}). Then, after some technical preliminaries
(\cref{sec:prelim}), we formalize the properties (\cref{sec:eager}, \cref{sec:lazy})
and describe testing (\cref{sec:testing}).

\section{Assumptions and Threat Model}
\label{sec:threat}

@@

Our properties are defined in terms of an abstract machine model that makes very few
assumptions about the machine that instantiates it.  We assume a register
machine with a program counter \(\PCname\) and stack pointer \(\rsp\); other
details of the machine's ISA are unimportant. We assume that we know ahead of time
the layout of memory, including code and which addresses are within
the bounds of the stack.

Our stack-safety concepts depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code (the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes).
%
We assume we are given a machine language program annotated with the locations of
instructions representing calls, returns, and function entry points. Although this
information would typically come from a compiler together with the machine code,
we do not assume anything about their provenance.

We further parameterize over the structure of the stack, particularly the means by
which a caller identifies the data that should be protected during a call. We term
this the {\em sealing convention} and model it as a relation between machine states
and addresses. Then from any call, those the sealing convention relates to the
state will be protected until the return.

This division of policy state from machine state is helpful when reasoning about
enforcement mechanisms that keep a significant amount of data separate from the
primary execution data to inform security. Obviously tag-based systems do this,
with tags and extra policy state separated from and untouchable by application code.
Another significant example is the shadow stack, where a second stack
mirrors the real one, and a return may only occur if both stacks agree on the
target address. Typically the shadow stack is isolated using a separate protection
mechanism.

The separation is less useful in protection mechanisms that interact more directly
with application code, such as Cheri-style capabilities. In these instances
the policy would not have auxiliary state, but would still refine the machine step
function to create a distinct failstop condition. Other techniques such as software-based
ones might never actually failstop, but fulfill the properties by never creating
traces that violate them.

\paragraph*{Threat Model}

Intuititively, stack safety means that a caller's stack frame is protected
from a confused or malicious callee's attempts to read or write its private data.
The attacker model therefore treats every caller as a defender and its callee as
an attacker, meaning that with nested calls, a single function may play both roles
at varying points.

We take a very strong threat model in which both caller and callee may execute
arbitrary machine code. This means that even the defender might be ill-behaved
to the point that it attempts to undermine its own protections, and we expect
enforcement mechanisms to ensure that it fails.
Hardware and timing attacks are out of scope, however.

@@

@@

Since we rely heavily on code annotations, we must ask: what if the annotations
are wrong? First, suppose a call or return occurs that is not annotated.
In this case, the answer is simple: as far as the properties are concerned, there has
been no call or return, only an instruction sequence resembling such an operation, and
therefore no associated change in permissions. This is particularly interesting in the
case of a false return, where control may appear to have returned to the caller, but
the enforcement mechanism must still protect the caller's memory.

@@

@@

%\paragraph*{Micro-Policies}
%
%Micro-Policies are a flexible tag-based, hardware-accelerated
%reference monitoring mechanism; they been applied to stack safety
%enforcement by \citet{DBLP:conf/sp/RoesslerD18}.
%Here the policy state consists of metadata tags (e.g., identifying stack frames)
%attached to each value in memory or registers, and the policy step
%function checks that each machine operation obeys a set of rules on tags
%(e.g., that the current $\PCname$ tag matches the tag on the stack
%location being accessed), and halts the machine if not.  Distinguishing
%ordinary machine state and policy state is very natural in this setting.
%% Here the machine state is extended with tags paired with all components of the
%% system, and a monitor checks for each operation that the tags on the operands
%% obey a set of rules. One could treat tags directly as part of the machine
%% state, and build the rules into the step function. But local state
%% encapsulation properties as we frame them are not intuitive in their
%% interaction with tags or other additional enforcement state, so it is helpful
%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
%% a software fail state.
%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
%%   policy? }

%\ifaftersubmission\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
%These approaches are represented as policies that
%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
%being produced by the compiler.) The set of policy states is again the unit type
%and the step function constant. Property enforcement comes from the fact that
%only ill-formed states can induce a trace that violates the property.
